Boosted Regression Trees for Model-Free Auto-Tuning

### 2011-12-06 paper/abstract.tex
m

\title{Boosted Regression Trees for Model-Free Auto-Tuning}

@@

\begin{abstract}

Autotuning is a widely used and effective technique for optimizing a
parametrized GPU code template for a particular type of computation on
particular hardware.
Its drawback is that exhaustive or even thorough autotuning can be very slow.
Furthermore, library abstraction boundaries provide operations such as image
filtering and matrix multiplication, which actually correspond to a large set
of potential problem configurations, with a potentially wide variety of memory
access patterns and computational bottlenecks.
How can we draw on data from previous auto-tuning of related problems on related hardware to make a
good decision on a novel problem?
This paper presents a machine learning approach to auto-tuning, in
which features of
(a) the current hardware platform,
(b) the kernel configuration
and (c) the problem instance are used to fit a regression model (boosted regression trees) that predicts
how much faster an implementation will be than a reference baseline.
Combinatorial optimization strategies that would normally be applied for autotuning can be applied to the regression model to find an implementation much more quickly.
We validate our approach using the filterbank correlation kernel described in [XXX].
We find that 0.05 seconds of a hill climbing on the regression model can achieve 80\% of the improvement brought by 120 seconds of real autotuning.
Our approach is not specific to filterbank correlation, or even to GPU autotuning: the approach of using a non-linear regression model on top of simple features applies to a variety of problem types, kernel types, and platforms.
A detailed model of the platform can provide useful features for cross-platform autotuning, but such a model is not necessary for same-platform autotuning.

\end{abstract}

>>>>

\title{Boosted Regression Trees for Data-Driven Auto-Tuning}
% Automatic
% of GPU Code
% Just-In-Time Code Specialization

@@

\begin{abstract}
Auto-tuning is a widely used and effective technique for optimizing a
parametrized GPU code template for a particular type of computation on
particular hardware.
Its drawback is that thorough or exhaustive auto-tuning requires compiling
many kernels, and calling each one many times; this process is slow.
Furthermore, library abstraction boundaries provide operations such as image
filtering and matrix multiplication, which actually correspond to a large set
of potential problem configurations with a wide variety of memory
access patterns and computational bottlenecks.
How can we draw on data from previous auto-tuning of related problems on related hardware to make a
good decision on a novel problem?
This paper presents a machine learning approach to auto-tuning, in
which features of
(a) the current hardware platform,
(b) the kernel configuration
and (c) the problem instance are used to fit a regression model (boosted regression trees) that predicts
how much faster an implementation will be than a reference baseline.
Combinatorial optimization strategies that would normally be applied for auto-tuning can be applied to the regression model to find an implementation much more quickly.
We validate our approach using the filterbank correlation kernel described in [anon].
We find that 0.05 seconds of hill climbing on the regression model can achieve
80\% of the improvement brought by 120 seconds of real auto-tuning.
Our approach is not specific to filterbank correlation, or even to GPU auto-tuning: the approach of using a non-linear regression model on top of simple features applies to a variety of problem types, kernel types, and platforms.
A detailed platform model can provide useful features for
cross-platform auto-tuning, but such a model is not necessary for same-platform autotuning.
\end{abstract}

### 2011-12-06 paper/draft1.tex
m

\title{Model-Based Auto-Tuning}

@@

\begin{abstract}
Each generation of GPU hardware presents unique programming challenges.
It is difficult to write software that obtains good performance on a range of hardware.

Auto-tuning is an effective technique for adapting a parametrized
GPU code template to current hardware.

Autotuning, however, is a relatively expensive kernel-selection mechanism - it
works by actually running the desired GPU computation in tens or hundreds of
ways before selecting the best implementation strategy for a particular
machine.

Futhermore library abstraction boundaries provide operations such as image
filtering and matrix multiplication, which actually correspond to a large set
of possible computations: for example the correlation of 1 large image with 1
small filters is  quite a different computational problem than the correlation
of 100 small images with 10 large filters.
This paper considers (XXX) strategies for using previous auto-tuning
measurements from similar problems to short-circuit computationally
expensive auto-tuning on a new problem of interest.

This paper presents a machine learning approach to generalized auto-tuning, in
which features of
(a) the current hardware platform,
(b) the kernel configuration
and (c) the problem instance are used to fit a regression model that predicts
how fast an implementation is (XXX).
We argue that autotuning using this surrogate model can be effective, while
being much faster than normal auto-tuning.

Whereas auto-tuning on problem XXX takes XXX seconds, our approach obtains
(XXX) of the speed in (XXX) of the time.

XXX: Hill-climbing (HC) search is effective alternative to grid.
XXX: How is HC search affected by mutation rate?
XXX: Purely random search is not an effective alternative to grid.

\end{abstract}

@@

\begin{figure}
\includegraphics[scale=.5]{speedup_295.pdf}
\caption{Speedup of various auto-tuning strategies over a good reference
implementation.
Each strategy was evaluated the same set of 83 problem configurations, which
was disjoint from the problem configurations used to build the model for
model-based hill-climbing.
The vertical dashed lines are positioned at the geometric mean speedup (XXX?)
of each strategy. The grid and hill-climbing approaches
tested 73 and 75 configurations respectively, and took 45?? XXX
seconds on average.
The model-based approach tested 0 configuration evaluations, and took 0.05
seconds on average, even with a naive Python implementation.
}
\label{fig:speedup_295}
\end{figure}

>>>>

\title{Boosted Regression Trees for Model-Free Auto-Tuning}

@@

\begin{abstract}
Autotuning is a widely used and effective technique for optimizing a
parametrized GPU code template for a particular type of computation on
particular hardware.
Its drawback is that thorough or exhaustive auto-tuning can be very slow.
Furthermore, library abstraction boundaries provide operations such as image
filtering and matrix multiplication, which actually correspond to a large set
of potential problem configurations with a wide variety of memory
access patterns and computational bottlenecks.
How can we draw on data from previous auto-tuning of related problems on related hardware to make a
good decision on a novel problem?
This paper presents a machine learning approach to auto-tuning, in
which features of
(a) the current hardware platform,
(b) the kernel configuration
and (c) the problem instance are used to fit a regression model (boosted regression trees) that predicts
how much faster an implementation will be than a reference baseline.
Combinatorial optimization strategies that would normally be applied for auto-tuning can be applied to the regression model to find an implementation much more quickly.
We validate our approach using the filterbank correlation kernel described in [anon].
We find that 0.05 seconds of hill climbing on the regression model can achieve
80\% of the improvement brought by 120 seconds of real auto-tuning.
Our approach is not specific to filterbank correlation, or even to GPU auto-tuning: the approach of using a non-linear regression model on top of simple features applies to a variety of problem types, kernel types, and platforms.
A detailed platform model can provide useful features for
cross-platform auto-tuning, but such a model is not necessary for same-platform autotuning.
\end{abstract}

@@

\begin{figure*}
\centering
\includegraphics[scale=.42]{speedup_295.pdf}
\includegraphics[scale=.42]{speedup_480.pdf}
\caption{Speedup of various auto-tuning strategies over a good reference
implementation.
Each strategy was evaluated the same set of 83 problem configurations, which
was disjoint from the problem configurations used to build the model for
model-based hill-climbing.
The vertical dashed lines are positioned at the geometric mean speedup (XXX?)
of each strategy. The grid and hill-climbing approaches
tested 73 and 75 configurations respectively, and took 45?? XXX
seconds on average.
The model-based approach tested 0 configuration evaluations, and took 0.05
seconds on average, even with a naive Python implementation.
}
\label{fig:speedup_295}
\end{figure*}
